Cybersecurity Education and Accreditation:\\A UK Case Study

### 2019-06-22 FIE2019-final.tex
Minor tweak to Acknow.

\title{Cybersecurity Is Too Important\\To Be Left To The Specialists}
%\title{What Do We Mean By Cybersecurity Education?}  Tom's original

% \author{\IEEEauthorblockN{Tom Crick}
% \IEEEauthorblockA{\textit{School of Education} \\
% \textit{Swansea University}\\
% Swansea, UK \\
% thomas.crick@swansea.ac.uk}
% \and
% \IEEEauthorblockN{James H. Davenport}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Bath}\\
% Bath, UK \\
% j.h.davenport@bath.ac.uk}
% \and
% \IEEEauthorblockN{Alastair Irons}
% \IEEEauthorblockA{\textit{Faculty of Technology} \\
% \textit{name of organization (of Aff.)}\\
% Sunderland University, Sunderland, UK\\
% alastair.irons@sunderland.ac.uk}
% \and
% \IEEEauthorblockN{Tom Prickett}
% \IEEEauthorblockA{\textit{Department of Computing and Information Sciences} \\
% \textit{name of organization (of Aff.)}\\
% Northumbria University, Newcastle upon Tyne, UK \\
% tom.prickett@northumbria.ac.uk}
% }
\ifanon
\author{Anonymous}
\else
% compact IEEE author format for more than three authors...
\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
    H. Davenport\IEEEauthorrefmark{2}, Alastair
    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
\IEEEauthorblockA{\IEEEauthorrefmark{2}Dept.~Computer Science, 
University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Dept.~Computer\&Information Sciences, 
Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
\fi
\maketitle

\def\IntroYear{2014}  % when BCS started asking for CyberSecurity

\begin{abstract}
% submitted abstract (100-500 word limit):
There are numerous facets to cybersecurity education, from theory to
practice, hardware and software, to social and technical (as well as
other important dimensions). A multitude of national and international
model curricula and recommendations - from national academies, learned
societies and even governments -- have been presented and discussed in
recent years, with varying levels of impact on policy and practice. In
this paper we attempt to address the key questions ``what should the
generalist computer scientist/engineer know'' and ``how well is that
being taught''. (Cyber)security -- if thought about at all -- has
historically been left to specialists, frequently viewed as a
masters-level discipline; or left for development in professional
practice, sometimes called ``information assurance'' in the UK and
Europe, which worked, even if not well, in a sequential process model.

For example, in the world of software and systems development, we have
seen the disruption of Agile (invented c.2001) and DevOps (invented
c.2009), but not necessarily through the lens of security; however,
more recently there has been a recognition of the need for change,
exemplified by the emergence of ``DevSecOps''. If security is not to
be left to the experts, then the generalist must know about it. Thus,
in the context of widespread international computer
science/engineering curriculum reform - both in compulsory, as well as
post-compulsory education - what does this trend mean more generally
for institutions and educators, and how do we teach it?

In this paper, we frame a key social, cultural and economic global
challenge of cybersecurity in computing and engineering education by
analysing the UK's national security, economic, and skills policy
context, shaping the practical considerations for education
professionals from schools through to universities (and into industry
and professional practice). Through this analysis of UK educational
policy and practice, we make a number of recommendations for future
cybersecurity educational initiatives and interventions, including a
number of innovative practical suggestions for educators and
curriculum designers, with a view for potential adoption and adaption
in other jurisdictions internationally.
\end{abstract}

\begin{IEEEkeywords}
Cybersecurity, accreditation, curricula, computer science education, UK
\end{IEEEkeywords}

@@

Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
It needs attention across the spectrum:
\begin{quote}
change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation. \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
\end{quote}

With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}. There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.  We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).

Nevertheless, it is one thing to write national curricula, specifications and requirements, and
another thing to deliver appropriate and relevant education and skills; furthermore, one could
reasonably ask how well this is done in practice.

\subsection{CyberSecurity: for all or for specialists?}
In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.

This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one. On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
% JHD what are these?  Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.

%AI: GDPR essentially expects CyberSecurity to be in all coding.

In particular the expectation of ``Privacy by Design'' and and by ``Privacy by Default'' has significant implications for all software development with a European context that processes personal data. Whilst both ``Privacy by Design'' and ``Privacy by Default'' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.

However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}. This further emphasizes the need for mainstream software developers and by extension Computer Science graduates to be versed in relevant security practices.

The Payments Industry \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code. This would seem to be particularly the case with ``agile'' processes \cite{Bartsch2011a}.

Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}. However, given the shortage of specialists, in practice a good generalist with some cybersecurity expertise can often get a ``specialist'' job.
\subsection{Research Questions}
There are various levels of specialism at which cybersecurity education can be addressed.
\begin{enumerate}
\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The general CS graduate.
\item The general CS masters graduate.
\item The specialist CS graduate.
\item The specialist CS masters graduate.
\end{enumerate}
The focus of this paper is on 2--3: the general CS graduate.
\begin{description}
\item[RQ1]What should be taught to the generalist, and how?
\item[RQ2]Should this be taught stand-alone or integrated? 
\item[RQ3]How might accreditation regard cybersecurity Education, or help with it?
\end{description}
\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
\begin{quote}
The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other Knowledge Areas.
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
\begin{quote}
Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
\end{quote}

But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and his research was based on web scraping, it represents a lower bound.

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}
\section{Policy Context}

% TC to add policy+skills context
% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en

There has been substantial CS/digital curriculum reform across the
UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
--- but what has happened to the focus on cybersecurity?  The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).

UK national economic skills priority
e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
(especially \cite{JCNSS2018a}
and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
\url{https://www.cybersecuritychallenge.org.uk/}).

There is large-scale media
attention on the ``cybersecurity skills gap''
e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
and
\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
etc -- link to UK Digital Strategy and UK Industrial Strategy.

\section{Challenges}

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
%{\bf JHD to research ACM general curriculum;}
The ACM general curriculum \cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).

Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.

%AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
  Should we be teaching it because it's underpinning? because there's a skill shortage?  JHD would argue both.

% should we frame this as one or more case studies?
% do we have any national survey data?

\subsection{PCI DSS}\label{sec:PCIDSS}
The Payments Card Industry Data Security Standards \cite{PCI2018b} underpin all processing of credit/debit cards. Nevertheless, they are very rarely mentioned in generalist computer scientist courses --- one payments industry person commented ``I've given up even asking if recruits have heard of PCI DSS: it's so rare''.

This would not matter so much if everyone handling payments data were sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS. Their response was that the customer wanted it!''.} PCI DSS course. However, the payments business is now so spread across websites, often run by SMEs, or non-specialists, that some provision should be made in courses.  Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.

%Alastair to write about their Credit Fraud experience (MSc cybersecurity).  HE Academy grant (to be referenced).

\subsection{Educational Resources: SQL Injection}\label{sec:SQL}
It is 15 years since \cite{Guimaraesetal2004} 
wrote ``All the topics listed above should be presented in the first
Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}. SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}). Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.  \cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
injection or script injection are known and documented.''.

Clearly such a major weakness should be well-taught. One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cybersecurity course.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.

>>>>

\title{Cybersecurity Education and Accreditation:\\A UK Case Study}
%\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
%\title{What Do We Mean By Cybersecurity Education?}  Tom's original

% \author{\IEEEauthorblockN{Tom Crick}
% \IEEEauthorblockA{\textit{School of Education} \\
% \textit{Swansea University}\\
% Swansea, UK \\
% thomas.crick@swansea.ac.uk}
% \and
% \IEEEauthorblockN{James H. Davenport}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Bath}\\
% Bath, UK \\
% j.h.davenport@bath.ac.uk}
% \and
% \IEEEauthorblockN{Alastair Irons}
% \IEEEauthorblockA{\textit{Faculty of Technology} \\
% \textit{name of organization (of Aff.)}\\
% Sunderland University, Sunderland, UK\\
% alastair.irons@sunderland.ac.uk}
% \and
% \IEEEauthorblockN{Tom Prickett}
% \IEEEauthorblockA{\textit{Department of Computing and Information Sciences} \\
% \textit{name of organization (of Aff.)}\\
% Northumbria University, Newcastle upon Tyne, UK \\
% tom.prickett@northumbria.ac.uk}
% }
% \ifanon
% \author{Anonymous}
% \else
% compact IEEE author format for more than three authors...
\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
    H. Davenport\IEEEauthorrefmark{2}, Alastair
    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
\IEEEauthorblockA{\IEEEauthorrefmark{1}%School of Education, 
Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
\IEEEauthorblockA{\IEEEauthorrefmark{2}%Dept.~Computer Science, 
University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{3}%Faculty of Technology, 
Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{4}%Dept.~Computer\&Information Sciences, 
Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
%\fi
\maketitle

\begin{abstract}
% submitted abstract (100-500 word limit):
This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets). A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice. This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?

Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper. Instead a case study-based research approach, primarily focusing on the UK, has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students. Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively? Through this UK case study, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.

%The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
\end{abstract}

\begin{IEEEkeywords}
Cybersecurity, curricula, accreditation, computer science education, public policy, UK
\end{IEEEkeywords}

@@

Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, from credit reference agencies~\cite{equifax2017}, online dating websites~\cite{ashleymadison2015}, national airlines~\cite{BritishAirways2018a}, telecoms companies~\cite{talktalk2016}, and even between sovereign governments~\cite{ncsc2018}. These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
\begin{quote}
``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
\end{quote}
\noindent
These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}. As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, \cite{Bristol2019a} is still a work in progress). An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}. As a result, new programmes entering this space are free to conceptualise, design, package and market their initiatives, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent or ambition of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.

With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}. This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}. There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report its full findings. The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.

Nevertheless, it is one thing to write national curricula, specifications and requirements, and
another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.

\section{Cybersecurity: for all or for specialists?}

In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}. This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.

This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK’s critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.

The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the computer science discipline for a number of years~\cite{Hentea2006,schneider2013}. There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}. There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}. The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees. Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs. More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity should be used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.

Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, should cybersecurity be formally taught in schools, as part of a compulsory education? While aspects of ``e-safety'' and principles of protecting personal data are clearly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}. The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country, primarily focused on university graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).

% have merged into previous section
%UK national economic skills priority
%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
%(especially \cite{JCNSS2018a}
%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
%\url{https://www.cybersecuritychallenge.org.uk/}). 
%There is large-scale media attention on the ``cybersecurity skills gap''
%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
%and
%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
%etc -- link to UK Digital Strategy and UK Industrial Strategy.

% need a better title for this!
% \subsection{Some Examples}

% To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.

% On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payments, has also led to greater demand for expertise in this area.
% % JHD what are these?  Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.

% %AI: GDPR essentially expects Cybersecurity to be in all coding. 
% In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications. However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
% % can we evidence this?

% A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}. This further emphasises the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.

% This can be further illustrated by the demands of particular sectors for secure software. Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process, even after developers are trained annually in secure coding practices \cite[6.5]{PCI2018b}, cannot be trusted to deliver secure code.

\section{Research Questions}
There are various levels of specialism at which cybersecurity education can be addressed.
\begin{enumerate}
\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The general computer science graduate.
\item The generalist computer science masters graduate.
\item The specialist computer science graduate.
\item The specialist computer science masters graduate.
\end{enumerate}
The focus of this paper is on 2-3: the general Computer Science graduate.
\begin{description}
\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science  students?
\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
\item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
\end{description}

\subsection{Research Approach}
A UK-focused case study-based approach is adopted in this research. As is common in case study-based research, many alternative case studies could have been chosen. The cases are illustrative rather than comprehensive in terms of the available case studies or challenges. The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.

The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations. This is followed by the evaluation of the policy context. Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.

In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated. Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.

Finally, in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK). As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.

\subsection{Current State of Play}
The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
\begin{quote}
{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other KAs}}
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, indicates
\begin{quote}
{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
\end{quote}

But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots''. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}

\section{Challenges: RQ1 and RQ2}

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
%{\bf JHD to research ACM general curriculum;}
The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).

%AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
  % Should we remove this?  This raises pedagogic questions which are beyond the scope of this research to resolve. Should we be teaching content because it's underpinning?; or because there's a skill shortage?

\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}). \cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.  There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.

The PCI DSS \cite{PCI2018b} is one such standard. PCI DSS underpins all processing of credit/debit cards. Nevertheless, they are very rarely mentioned in generalist computer scientist courses. This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course. However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.  Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
